Software Citation Principles

### 2016-07-14 sc-principles/rebuttal/rebuttal.tex

%to highlight text
\usepackage{soul}
\usepackage[usenames, dvipsnames]{xcolor}
\newcommand{\hly}[1]{{\sethlcolor{yellow}\hl{#1}}}
\newcommand{\hlb}[1]{{\sethlcolor{SkyBlue}\hl{#1}}}
\newcommand{\hlg}[1]{{\sethlcolor{green}\hl{#1}}}

\usepackage[sort&compress]{natbib}
\bibpunct{[}{]}{,}{n}{}{;}
\usepackage[linkcolor=blue, citecolor=blue, colorlinks=true, breaklinks=true]{hyperref}

\usepackage{url,breakurl}

\newcommand{\katznote}[1]{ {\textcolor{blue} { ***DSK: #1 }}} % Dan
\newcommand{\niemnote}[1]{ {\textcolor{orange} { ***KEN: #1 }}} % Kyle
\newcommand{\asnote}[1]{ {\textcolor{red} { ***AMS: #1 }}} %Arfon

@@

\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\setlength{\parindent}{0in}
\setlength{\parskip}{0.4cm}

%%% BEGIN DOCUMENT
\begin{document}

We thank the editor and reviewers for their helpful comments, and have addressed their requests and questions here.
Specific changes or additions can be found in the marked revision.
% in response to the Editor, Reviewer \#1, and Reviewer \#2 are respectively highlighted in the marked manuscript in \hly{yellow}, \hlb{blue}, and \hlg{green}.

@@

\emph{First of all, I would like to thank the reviewers, who have done a fantastic job and provided very insightful comments for improving the paper. I hope that the authors will take them into serious consideration in the revision.}

\emph{From my personal perspective, this is a very good contribution, which addresses one of the most important topics in the ``citation'' domain, i.e. software citation. I totally agree with the reviewers about the readability of the paper, and I think this paper will deserve publication in PeerJ CS.}

\emph{However, there are some aspects that should be clarified in the revision. In particular:}

\begin{enumerate}
\item  \textbf{Comment:}
%\textbf{1. Comment:}
\emph{I would like to see a new section that describes the presentation of these principles at FORCE 2016, and includes also a summary of the most important outcomes derived from the discussion the authors have done during the conference. Did such outcomes change some of the principles and/or other claims in this paper?}

@@

\textbf{Our response:}

\item \textbf{Comment:}
\emph{A clear research question should be announced explicitly from the very beginning, and accompanied by a text that explains how the authors would like to address it.}

\textbf{Our response:}

\item \textbf{Comment:}
\emph{I suggest to include one or more appendixes with all the details about the use cases presented. This would make the paper more self-contained.}

\textbf{Our response:}
\niemnote{I will add the full use cases text as an appendix}

\item \textbf{Comment:}
\emph{The reviewers did not agree on some aspects described in the paper, e.g. "if the software authors ask that a paper should be cited, that should typically be respected". I would like to see an appropriate discussion on these points, so as to clearly address these disagreements.}

@@

\textbf{Our response:}

\item \textbf{Reviewer comment:}
\emph{This paper doesn't have an experimental design in the strict sense, as it is more of a position paper than a research article. The PeerJ guidelines state that "the submission should clearly define the research question", which is currently not the case for this paper. However, I see a research question between the lines, which could be phrased as "how can citations be applied to software products to account for their increasing importance in scientific research?". By making such a research question explicit, the PeerJ requirements can be met.}

\textbf{Our response:}

\item \textbf{Reviewer comment:}
\emph{The use cases described in Section 3 form the main method to answer the implicit research question. Given the importance of these use cases for this paper, I think reference [18] with the details about these use cases should be published in a more official way, either on a site like ArXiv or FigShare, or maybe even as supplemental material or appendix for this paper. I think having these use cases in the appendix of this paper would be the best solution, which would also increase the impact of these use cases and would strengthen the paper in general.}

\textbf{Our response:}
\niemnote{I will add the full use cases text as an appendix}

\item \textbf{Reviewer comment:}
\emph{As noted above, the logical structure of the paper is unclear, and therefore it is not clear what exactly the main findings are. I think the main findings should be the principles stated in the beginning, and the use cases would then form the main method to arrive at these findings. Alternatively, one could label the principles as "assumptions" and apply the use case study to validate these assumptions. In any case, I think the findings are valid but they should be labeled as preliminary, as it is not possible to arrive at a final conclusion for such kinds of questions. Only the future can tell whether software citations really work out.}

@@

\textbf{Our response:}

\item \textbf{Reviewer comment:}
\emph{line 88: ``which will be presented ...'' > ``which was presented ...''}

\textbf{Our response:}
We agree, and have made this change.

\item \textbf{Reviewer comment:}
\emph{line 89: ``We expect that this discussion may lead to a second, final version'': So this paper is not the final version? Please clarify.}

\textbf{Our response:}
Thank you for pointing this error out---we have corrected this statement to indicate that the past discussion led to a second version (what was submitted).

\item \textbf{Reviewer comment:}
\emph{line 97: ``the need for credit for application software underscores the need to overhaul the system of credit for all research products.'': Very true and very important. But then maybe the concept of a ``citation'' is obsolete altogether, and we shouldn't be advocating ``software citations'', but something new altogether? Some more discussion on this would be valuable, I think. This relates to my comment above about a long-term vision.}

@@

\textbf{Our response:}

\item \textbf{Reviewer comment:}
\emph{In the discussion about Unique identification (lines 292+), I miss the discussion of hash values as version identifiers, as applied by Git for instance. As the first author is affiliated with GitHub, I assume this aspect was left out intentionally, but why? As basically all software is nowadays maintained with version control systems such as Git, these identifiers come for free and are by design unique. So, I would think they are an important aspect to discuss for software citations.}

\textbf{Our response:}

\end{enumerate}

\begin{enumerate}

\item \textbf{Reviewer comment:}
\emph{The two tables are relevant and helpful, though I did have a slight problem with the legibility of my printed copy of Table 2 (page 5). The distinction between solid and open circles was not immediately clear to me, and I might suggest using a slightly larger version of these symbols.}

\textbf{Our response:}
We agree, and have changed the open symbol to a plus sign to improve legibility.

\item \textbf{Reviewer comment:}
\emph{The paper does not present a specific research finding, and as such does not feature a research question and experimental procedure that are expected of such papers. However, I think it makes in important contribution to the acceptance of software as part of the process and output of research activity, and reflects a consensus formed among a significant community of researchers. As such, I believe it justifies a place in the formally published literature.}

\textbf{Our response:}
We agree with the reviewer's description of our paper, and appreciate the view that it should be formally published.

\item \textbf{Reviewer comment:}
\emph{Lines 32--35: as a justification for treating software differently from data, I found this passage raised more questions than answers, and that the comments here were uncompelling and unsupported. For example, some published data is updated on a frequency comparable with some software (e.g. the bi-monthly release schedule of FlyBase is comparable with many software projects - \url{http://flybase.org/static_pages/docs/release_schedule.html}). Also, I think the distinction between data and software is less clear-cut than the authors suggest (e.g. is a schema or stylesheet data or software?). I would see the difference between software and data as more akin to the difference between research methods and research findings, but that is just a personal opinion. My suggestion would be that this justification is not really necessary to the overall purpose of the paper, and might be dropped without significant loss of value.}

\item \textbf{Our response:}

\item \textbf{Reviewer comment:}
\emph{Line 57, et seq: [Nit] I would find this easier to read if the sentence introducing the list of reasons were moved to the same page as the list itself.}

\textbf{Our response:}
\niemnote{The formatting will change in the final published paper, so this might not need to be addressed. Or, we can bring in some of the commented statements from above the table}

\item \textbf{Reviewer comment:}
\emph{Line 88: ``will be presented'' refers to an event now in the past: change to ``was presented''?}

\textbf{Our response:}
We agree, and have made this change.

\item \textbf{Reviewer comment:}
\emph{Line 113: [Nit] ``work using software'' appears to be very broadly scoped, and could apply to just about any activity using a computer. Suggest ``research using software''.}

\textbf{Our response:}
We agree, and have made this change.

\item \textbf{Reviewer comment:}
\emph{Table 2, page 5: I had a problem with the legibility of my printed copy of this table. The distinction between solid and open circles was not immediately clear to me. I might suggest using a slightly larger version of these symbols, or maybe more distinct symbol shapes.}

\textbf{Our response:}
We agree, and have replaced the open bullet with a plus sign for better legibility.

\item \textbf{Reviewer comment:}
\emph{Table 2, page 5: [Comment] This table contains a lot of information, and much food for debate. I was surprised by some omissions, which may mean that I wasn't fully understanding the concepts being presented:\\
\\
(a) I find it hard to imagine any of the use cases not benefiting from availability of a ``Description'', especially 1, 2, 9, and 16.\\
(b) I would have expected ``Indexed citations'' to be useful for 5: getting credit.\\
(c) For publishing a software paper, I would expect all of the basic requirements other than ``Indexed citations'' to be at least beneficial.\\
(d) For 12: build a software catalog/registry, would there not be a role for indexed citations?\\
(e) For 16: Store software try, I'd expect pretty much all of the Basic requirements other than ``Indexed citations'' to be relevant, and particularly ``Description'' and ``Keywords''.\\
(f) I think there is one use-case missing, which I think is commonly faced by researchers: viz. ``Evaluate suitability of software for a given task'' - I think this is distinct from discovery, and typically involves making a choice between different available software packages, or to use an existing package, modify an existing package or develop something new.}

@@

\item \textbf{Reviewer comment:}
\emph{Table 2, page 5: [Comment] This table contains a lot of information, and much food for debate. I was surprised by some omissions, which may mean that I wasn't fully understanding the concepts being presented:\\
\\
(a) I find it hard to imagine any of the use cases not benefiting from availability of a ``Description'', especially 1, 2, 9, and 16.\\
(b) I would have expected ``Indexed citations'' to be useful for 5: getting credit.\\
(c) For publishing a software paper, I would expect all of the basic requirements other than ``Indexed citations'' to be at least beneficial.\\
(d) For 12: build a software catalog/registry, would there not be a role for indexed citations?\\
(e) For 16: Store software try, I'd expect pretty much all of the Basic requirements other than ``Indexed citations'' to be relevant, and particularly ``Description'' and ``Keywords''.\\
(f) I think there is one use-case missing, which I think is commonly faced by researchers: viz. ``Evaluate suitability of software for a given task'' - I think this is distinct from discovery, and typically involves making a choice between different available software packages, or to use an existing package, modify an existing package or develop something new.}

\textbf{Our response:}

\item \textbf{Reviewer comment:}
\emph{Line 175: [Nit] ``in(1)'' Missing space?}

\textbf{Our response:}
This has been fixed.

\item \textbf{Reviewer comment:}
\emph{Lines 241--242: ``Similarly, the software metadata that is recorded as part of data provenance should be a superset of the metadata recorded as part of software citation.'' I found this non sequitur (or am not seeing the logical connection). I can imagine reasons to cite software that does not form part of the provenance chain of results presented (e.g. alternatives considered).}

@@

\textbf{Our response:}

\item \textbf{Reviewer comment:}
\emph{Lines 245--247: ``In general, we intend the software citation principles to cover the minimum of what is necessary for software citation for the purpose of software identification. Other use cases (e.g., provenance, reproducibility) may lead to additional requirements (i.e., enhanced metadata).'' This statement feels to me to be at odds with table 2, which presents ``Use cases and basic metadata requirements for software citation'', which includes ``reproducibility'' as a use-case. This leaves me wondering if the use cases in table 2 are going beyond what is strictly needed for the purposes of software citation.}

\textbf{Our response:}

\item \textbf{Reviewer comment:}
\emph{Line 327--328: ``When software exists as both source code and another type, we recommend that the source code be cited''. The term ``exists'' here doesn't seem quite right to me. E.g., proprietary software may ``exist'' as source code, but be unavailable. Suggest ``exists and is accessible'' or just ``is accessible''?}

\textbf{Our response:}
Agreed; this change has been been.

\item \textbf{Reviewer comment:}
\emph{Line 348: ``Updates to this document''. This paper is being presented for consideration as a part of the formal academic record - as such, I don't think ``this document'' is something which may be updated. Suggest ``Updates to these principles''.}

\textbf{Our response:}
Agreed; this change has been been.

\end{enumerate}

### 2016-07-15 sc-principles/software-citation-principles.tex
added appendix with use cases from Google Doc; re editor #4 & reviewer 1 #4

>>>>

\section{Software citation use cases}
\label{app:usecases}

This appendix records an edited, extended description of the use cases discussed in~\S\ref{sec:use_cases}, originally found in~\cite{SC-Use-Cases}.

\subsection{Researcher who uses someone else's software for a paper}

One of the most common use cases may be researchers who use someone else's software and want to cite it in a technical paper. 
This will be similar to existing practices for citing research artifacts in papers.

``Requirements'' for researcher:
\begin{itemize}
\item Name of software
\item Names of software authors/contributors
\item Software version number and release date, or download date
\item Location/repository, or contact name/email (if not publicly available)
\item Citable DOI of software
\item Format for citing software in text and in bibliography
\end{itemize}

Possible steps:
\begin{enumerate} \setcounter{enumi}{0}
\item Software developers create CITATION file and associate with source code release/repository.
\item Researcher finds and uses software for research paper.
\item Researcher identifies citation metadata file (e.g., ``CITATION'' file) associated with downloaded/installed software source code or in online repository/published location. CITATION file includes necessary citation metadata. CITATION file may include BibTeX entry, suggested citation format
\item Researcher cites software appropriately, e.g. in methodology section; reference included in bibliography.
\end{enumerate}

\subsection{Researcher who uses someone else's software for new software}

In this case, a researcher develops new software that incorporates or depends on existing software. 
In order to credit the developer(s), the researcher will include citations in his/her source code, documentation, or other metadata in a similar manner to papers

Requirements for researcher:
\begin{itemize}
\item Name of software
\item Names of software authors/contributors
\item Software version number and release date
\item Location/repository
\item Citable DOI of software
\item Format for citing software in source code, documentation, or citation metadata file
\end{itemize}

Possible steps:
\begin{enumerate} \setcounter{enumi}{0}
\item Assume that software developers have created a CITATION file and associated with the source code release/repository.
\item Researcher finds and uses software in the development of new software.
\item Researcher identifies citation metadata file (e.g., ''CITATION'' file) associated with downloaded/installed software source code or in online repository/published location. CITATION file includes necessary citation metadata. CITATION file may include BibTeX entry, suggested citation format.
\item Researcher cites software in source code, documentation, or other metadata-containing file.
\end{enumerate}

\subsection{Researcher who contributes to someone else's software (open source project)}

maybe a variant of above\niemnote{add more to this?}

\subsection{Researcher who wants to know who uses the researcher's software}

This case is similar to a researcher who wants to find other papers/publications that cite a particular paper.
A researcher wants to gauge the usage of her software within or across communities and measure its impact on research for both credit and funding.

Requirements:
\begin{itemize}
\item Uniquely identify software
\item Indexed citations of software 
\item Indexed papers that use software
\end{itemize}

Steps:
\begin{enumerate}
\item Researcher finds software official name or unique DOI in metadata associated with downloaded/installed source code or in online repository/published location.
\item Researcher searches for software, may use online indexer (e.g., Scopus, Web of Science, Google Scholar) using software name or DOI.
\item Online indexer presents entry for software with list of citations, if any.
Ideally, entry will also include metadata contained in software CITATION file and citation example.
\end{enumerate}

\subsection{Researcher who wants to publish about a piece of software}

The research wants to publish about a version of software they have produced.
A key part of this use case is to be able to connect the given narrative to a specific version of the software in questions and connect that in large story.

Requirements:
\begin{itemize}
\item Name of software
\item Names of software authors/contributors
\item Location/repository
\item Citable DOI of Software
\item Links to older versions of software
\end{itemize}

\subsection{Publisher who wants to publish papers that cite software}

\niemnote{add description?}

Requirements for publisher:
\begin{itemize}
\item Name of software
\item Names of software authors/contributors
\item Location/repository
\item Citable DOI of software
\item Format for citing software in, e.g., JATS, as well as references in the text itself
\end{itemize}

\subsection{Indexer (e.g., Scopus, WoS, Scholar, MS Academic Search) who wants to build a catalog of software}

Provide an index over the software that is used within the research domain. Track how that software is being used by different groups of researchers and to what ends.

Requirements:
\begin{itemize}
\item Uniquely identify pieces of software used by the research literature
\item Connect authors and organizations to that software
\item Connect various software versions together
\end{itemize}

\subsection{Domain group (e.g., ASCL, bioCADDIE), Libraries, and Archives (e.g., University library, laboratory archive, etc.) wants to build a catalog\slash registry of institutional or domain software}

There are two different examples here: One is building a catalog/archive of software produced by those affiliated with the institution.
The other is along the lines of Sayeed Choudhury's note that ``data are the new special collections.''
An institution may choose to build a catalog/archive of many things within a single topic or subject in order to secure all the software on a certain topic or build a collection that may draw users to their establishment, much like special collections now do for university libraries and archives.

\subsection{Repository showing scientific impact of holdings}

A repository that archives and/or maintains a collection of software.
The repository would like to address usage and impact of software in its holding.
Usage would aid potential users whether the software is being actively maintained or developed or has been superseded.
Both would help repository know how to direct resources, e.g., maintenance, training etc.

Requirements:
\begin{itemize}
\item Code name ideally, a unique identifier
\item Relationships to previous versions
\item Connect to repository
\item Connect to research
\end{itemize}

similar to Funder case

mineable via repository name, code name, DOI, or citation.\niemnote{complete?}

\subsection{Funder who wants to know how software they funded has been used}

steps/requirements\niemnote{add more to this?}

\subsection{Researcher gets credit for software development at the academic\slash governmental institution, in professional career, etc.}

\niemnote{add description?}

Requirements for researcher:
\begin{itemize}
\item Name of software
\item Names of software authors/contributors
\item Location/repository
\item Citable DOI of software
\item Format for citing software in an official CV, in a departmental/institutional review report, etc.
\item Role in the software creation, that is linked to version or component
\item Role in contributing to the software as a ``package'' (not just lines of code) development of benchmarks, testing, documentation, tutorials etc.
\end{itemize}

\subsection{Researcher who wants to ``reproduce'' another person\slash group's analysis}

\niemnote{add description?}

Requirements for researcher:
\begin{itemize}
\item Name of software
\item Location/repository for the exact release that was used 
\item DOI or other persistent handle for that specific release
\item Release has all components necessary for reproducing the work (note that this ideally also means sample inputs and outputs)
\end{itemize}

\subsection{Researcher who benchmarks someone else's software with or without modification on one or many hardware platforms for publication}

\niemnote{add description?}

Requirements for researcher:
\begin{itemize}
\item Name of software
\item Names of software authors/contributors
\item Software version number and release date
\item Location/repository
\item Citable DOI of software or paper recommended for citation
\item Format for citing software in source code or citation metadata file
\end{itemize}

Possible steps:
\begin{enumerate} \setcounter{enumi}{0}
\item Software developers create CITATION file and associate with source code release/repository.
\item Researcher finds and uses software in the development of new software.
\item Researcher identifies citation metadata file (e.g., CITATION file) associated with downloaded/installed software source code or in online repository/published location. CITATION file includes necessary citation metadata. CITATION file may include BibTeX entry, suggested citation format.
\item Researcher cites software in source code, documentation, or other metadata-containing file.
\end{enumerate}

\subsection{Reference management system used by researchers to author a manuscript}

\niemnote{add description?}

Requirements for reference manager:
\begin{itemize}
\item Names of software authors/contributors
\item Software version number and release date
\item Location/repository
\item Citable DOI of software or paper recommended for citation
\item Format for citing software in citation metadata file
\item Citation metadata tags embedded in DOI landing page/software project page for easy ingest
\end{itemize}

Possible steps:
\begin{enumerate}
\item Reference management system such as EndNote, Mendeley, Zotero, etc. builds affordances for software references.
\item Researcher finds software citation and adds it to their reference manager library, by (a) importing from the CITATION file (e.g., BibTeX, RIS), or (b) clicking on, e.g., an ``add to Zotero library'' widget in web browser.
\item Researcher writes a paper and uses the reference manager to generate citations or bibliography.
\end{enumerate}

\subsection{Researcher who wants to find a piece of software to implement a task}

This is the case where a research is looking for software to use but wants to understand whether it is being used in a scholarly fashion.
For example, a researcher searches through a software repository and finds a package that might be useful.
They look to find whether it has been used by others in the scientific literature.

Requirements
\begin{itemize}
\item Either the software documentation page has a reference to existing literature that makes use of it. 
\item There is a mechanism to look it up.
\end{itemize}

\subsection{Researcher wants to record the software that generated some data}

This is the case where a researcher is using some software to perform an analysis, either of a physical sample or of data.
The researcher needs to know which version was used, for example in case a bug was fixed.
Note that knowing the software and its version is not sufficient to determine the ``conditions'' of the analysis, but they are essential.

Requirement:
The analysis, or the generated data, has information about the software used.

\subsection{Researcher who wants to reproduce experience of use of a particular software implementation in context}

\niemnote{This one was not adopted into table; should we keep it, remote here, or somehow indicate why it was not adopted?}

Researcher is engaged in historical/cultural research -- e.g., study of video games as cultural artifacts.

Requirements
\begin{itemize}
\item Name of software
\item Software version number
\item Documentation of the execution environment/context
\item Location/repository for virtual machine (or equivalent) comprising both software and execution environment/context
\item Persistent identifier associated with virtual machine instance (or equivalent) comprising both software and execution environment/context
\end{itemize}

Possible steps
\begin{enumerate}
\item Researcher obtains persistent ID from citation
\item Research uses a persistent ID resolution service to resolve ID to a location of  an executable VM instance in a repository 
\item Researcher obtains VM in the repository, executes it, and interacts with software 
\end{enumerate}

%References:
%Use case examples and commentary
%Library of congress report on software preservation:
%http://www.digitalpreservation.gov/meetings/preservingsoftware2013.html
%CASPAR Report on Software Significant Properties
%http://sigsoft.dcc.rl.ac.uk/twiki/pub/Main/SigSoftTalks/SignificantPropertiesofSoftware-r1.doc
%Examples of VM archiving and repositories
%https://olivearchive.org/
%https://archive.org/details/internetarcade
